No Detectable Evidence of Genetic Associations with Anadromous and Resident Masu Salmon Sampled within Hokkaido

### 2022-08-31 paper/paper_draft.Rmd
updated draft

Due to the history of hydrologic modifications such as the development of reservoirs, dams, and irrigation systems, there has been an apparent decrease in the number of migratory fishes in recent years [@Pringle2000]. In North America, for example,there have been significant decreases in diversity within pacific salmonid species due to hydrologic and dam alterations, and atlantic salmon and anadromous species of sturgeons are also impacted negatively because of a loss of portions to their habitats upstream of dams [@Pringle2000]. Overall, the diversity within migratory fish species have largely declined by an average of 70% between 1970 and 2016 due to these dams blocking the fishes' migratory paths [@deinet2020living].

@@

The decision to migrate in masu salmon is thought to be primarily influenced by environmental factors such as temperature alongside other factors such as the salmon's body size, growth rate, lipid stores, and is heavily sex influenced [@Morita2014; @Arostegui2019]. Consequently, changes in environmental factors, such as those caused by climate change, have been known to impact the decision of the masu to be either resident or anadromous [@Morita2014]. For example, Morita et al (2014) has shown that the number of resident males have increased while the proportion of anadromous males and the proportion of delayed migrants within both females and males have decreased as the temperatures have increased, which has led to a significant decrease in the number of anadromous individuals overall.

In contrast, the genetic factors that contribute to migration in masu salmon have not been well studied.  Despite the prevalence of environmental control of migration in masu, it is not unreasonable to expect genetic elements to play some role in governing migratory decisions in the species. For example, within steelhead, there is strong evidence suggesting both an environmental and genetic basis for migratory life history. In 1944, Naeve conducted common garden and transplant experiments with sympatric rainbow trout and steelhead and found that migratory status was heritable after observing that rainbow trout offspring were considerably more likely to be resident themselves, whereas  crosses between anadromous individuals resulted in the most anadromous smolts. It is not surprising, therefore, that a recent study by Hess et al. [-@Hess2016], on 15,239 single nucleotide polymorphisms (SNPs) found of three SNPs from a gene that was strongly linked with adult migration timing, and that 18 total SNPs that were connected to 60% of the trait variation. Despite this, however, growth rates and body condition are known to influence anadromy in the species—individuals which grow faster as juveniles are more likely to stay as residents than otherwise [@Kendall2015]. Growth rates are also known to be influenced by large genetic elements that differ between anadromous and resident individuals, implying that genetics may contribute to migration indirectly as well [@MILLER2012]. Taken together, it seems clear that migration in steelhead is  jointly controlled by the interplay of genetic and environmental factors.

@@

```{r, include=FALSE, eval=TRUE}
dat<-read.table("../data/genotypes.geno")                        
meta<- read.table("../data/sample_meta.txt", header=TRUE)
snp.meta<- dat[,1:2]
colnames(snp.meta)<- c("chr", "position")
library(snpR)
# dat<- import.snpR.data(dat[,-c(1:2)], snp.meta = snp.meta, sample.meta = meta)
# dat<- filter_snps(dat, maf = 0.05, hwe = 0.000001)
# number_snps_after_filtering <- nsnps(dat)
# 
# #calculating genetic diversity
# dat <- calc_ho(dat, c("pop", "phen", "pop.phen"))
# get.snpR.stats(dat, c("phen", "pop","pop.phen"), stats="ho") $weighted.means
# get.snpR.stats(dat, c("pop", "phen", "pop.phen"))
# dat <- calc_fis(dat, c("pop", "phen"))
# dat <- calc_pi(dat, c("pop", "phen"))
# get.snpR.stats(dat, c("pop"), stats="pi") $weighted.means
# dat <- calc_private(dat, c("phen", "pop","pop.phen"))
# get.snpR.stats(dat, c("pop"), stats="private") $weighted.means
# get.snpR.stats(dat, c("pop", "phen", "pop.phen"))
# dat <- calc_pairwise_fst(dat, c("phen", "pop","pop.phen"), boot = 100, boot_par = 2)
# get.snpR.stats(dat, c("phen", "pop","pop.phen"), stats="fst") $weighted.means
# get.snpR.stats(dat, c("pop", "phen"), stats= c("ho", "pi", "fst")) $weighted.means
# 
# gp<- run_genomic_prediction(dat, response = "phen", iterations = 10000, burn_in = 1000, thin = 100)
# library(data.table)
# cv<- cross_validate_genomic_prediction(dat, response = "phen", iterations = 10000, burn_in = 1000, thin = 100)

@@

We also conducted a Principle Component Analysis chart (PCA) using the ANGSD-IBS method [@Korneliussen2014]. To look for patterns in relatedness between resident and andromous individuals, we calculated the mean PC1 and PC2 coordinates for each stream, then calculated the distance between each individual sample the mean for the other population. We fit a simple linear regression with the distance to the other population cluster as the response variable and the migratory phenotype as the explanatory variable.

>>>>

Due to the history of hydrologic modifications such as the development of reservoirs, dams, and irrigation systems, there has been an apparent decrease in the number of migratory fishes in recent years [@Pringle2000]. In North America, for example,there have been significant decreases in diversity within pacific salmonid species due to hydrologic and dam alterations, and atlantic salmon and anadromous species of sturgeons are also impacted negatively because of a loss of portions to their habitats upstream of dams [@Pringle2000]. Overall, the diversity within migratory fish species have largely declined by an average of 70% between 1970 and 2016 due to dams and other factors  [@deinet2020living].

@@

The decision to migrate in masu salmon is thought to be primarily influenced by environmental factors such as temperature alongside other factors such as the salmon's body size, growth rate, lipid stores, and is heavily sex influenced [@Morita2014; @Arostegui2019]. Migration and residency in the closely related steelhead salmon (Oncorhynchus mykiss) are controlled by similar factors. For example, resident steelhead tend to be larger within the first few months of life when compared to anadromous individuals (Morita et al., 2014). Of course, differences in environmental factors can influence growth rates and therefore also migration in steelhead. For example, during poor nutrient conditions, a higher proportion of individuals than expected from genetics may be anadromous; whereas during flooding or nutrient rich periods, a higher proportion of resident individuals may be observed [@Kelson2020]. Regardless of environmental conditions, however, those individuals within a population which grow faster as juveniles are more likely to stay as residents than otherwise, even under similar conditions [@Kendall2015].

Migration comes with a tradeoff. In steelhead, migratory juveniles tend to have low survival rates to adulthood, which is  compensated for by the high fecundity of anadromous females that can significantly exceed that of residents [@Pearse2019]. This pattern likely holds true for masu as well (Morita et al., 2014). This trade-off is not as pronounced in male steelhead, which can mature as early as freshwater residents, avoiding the high mortality associated with migration [@Pearse2019]. This results in greater frequency of anadromy in females; a trend also seen in masu salmon (Morita et al., 2014). Consequently, changes in environmental factors, such as those caused by climate change, have been known to impact the decision of the masu to be either resident or anadromous [@Morita2014]. For example, Morita et al (2014) has shown that the number of resident males have increased while the proportion of anadromous males and the proportion of delayed migrants within both females and males have decreased as temperatures have increased over time, leading to a significant decrease in the overall number of anadromous individuals.

In contrast to environmental factors, the genetic factors that contribute to migration in masu salmon have not been well studied. Despite the prevalence of environmental control of migration in masu, it is not unreasonable to expect genetic elements to play some role in governing migratory decisions in the species. For example, within steelhead, there is strong evidence suggesting both an environmental and genetic basis for migratory life history. In 1944, Naeve conducted common garden and transplant experiments with sympatric rainbow trout and steelhead and found that migratory status was heritable after observing that rainbow trout offspring were considerably more likely to be resident themselves, whereas crosses between anadromous individuals resulted in the most anadromous smolts. Indeed, recent studies have found that  the decision to migrate in steelhead is strongly linked to a large genomic inversion on chromosome Omy5 throughout their geographic range which seems to maintain a set of linked genes that controls migratory behavior [@Arostegui2019; @Kelson2020; @Pearse2014). The genetic basis for migration is also likely indirectly mediated by physiological characteristics such as growth rate, which are also known to be influenced by the same large genetic elements that differ between anadromous and resident individuals [@Miller2021]. Taken together, it seems clear that migration in steelhead is jointly controlled by the interplay of genetic and environmental factors.

@@

```{r, include=FALSE, eval=TRUE}
dat<-read.table("../data/genotypes.geno")                        
meta<- read.table("../data/sample_meta.txt", header=TRUE)
snp.meta<- dat[,1:2]
colnames(snp.meta)<- c("chr", "position")
library(snpR)
dat<- import.snpR.data(dat[,-c(1:2)], snp.meta = snp.meta, sample.meta = meta)
dat<- filter_snps(dat, maf = 0.05, hwe = 0.000001)
number_snps_after_filtering <- nsnps(dat)

#calculating genetic diversity
#dat <- calc_ho(dat, c("pop", "phen", "pop.phen"))
#get.snpR.stats(dat, c("phen", "pop","pop.phen"), stats="ho") $weighted.means
#get.snpR.stats(dat, c("pop", "phen", "pop.phen"))
#dat <- calc_fis(dat, c("pop", "phen"))
#dat <- calc_pi(dat, c("pop", "phen"))
#get.snpR.stats(dat, c("pop"), stats="pi") $weighted.means
#dat <- calc_private(dat, c("phen", "pop","pop.phen"))
#get.snpR.stats(dat, c("pop"), stats="private") $weighted.means
#get.snpR.stats(dat, c("pop", "phen", "pop.phen"))
#dat <- calc_pairwise_fst(dat, c("phen", "pop","pop.phen"), boot = 100, boot_par = 2)
#get.snpR.stats(dat, c("phen", "pop","pop.phen"), stats="fst") $weighted.means
#get.snpR.stats(dat, c("pop", "phen"), stats= c("ho", "pi", "fst")) $weighted.means
# 
# gp<- run_genomic_prediction(dat, response = "phen", iterations = 10000, burn_in = 1000, thin = 100)
# library(data.table)
# cv<- cross_validate_genomic_prediction(dat, response = "phen", iterations = 10000, burn_in = 1000, thin = 100)

@@

We also conducted a Principle Component Analysis chart (PCA) using the ANGSD-IBS method [@Korneliussen2014]. To look for patterns in relatedness between resident and anadromous individuals, we calculated the mean PC1 and PC2 coordinates for each stream, then calculated the distance between each individual sample the mean for the other population. We fit a simple linear regression with the distance to the other population cluster as the response variable and the migratory phenotype as the explanatory variable.
